The Minimal Retroreflective Microfacet Model

### 2025-12-23 article.tex
Fix hyphenation

\teaser{
  \vspace*{-0.5cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/retro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
  \caption{Renders of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row). The lighting in all cases is aligned with the view direction, exhibiting a strong retro-reflective peak in the MRM model.}
   \vspace{-0.5cm}
  \label{fig:teaser}
}

@@

Rendering of retro-reflective materials is useful in a variety of contexts. For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some example renders). Realistic rendering of such materials is important for design and visualization purposes, and in a predictive rendering context for example for simulation of visibility in training of autonomous vehicles or for safety certification of road markings.

@@

\paragraph{Previous works.}
Modeling of retro-reflectivity in the surface scattering function has received relatively little attention so far. For example, \citet{Guo2017Retro} proposed a model to incorporate the retro-reflectivity of prismatic sheets. However, their model only provides pure retro-reflection which would lack artistic control. They later extended this model to account for imperfect retro-reflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}. However, due to the coupling of retro-reflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model. \citet{Hericz2016} proposed a retro-reflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.

@@

For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving and reciprocal, c) efficient and easy to implement, ideally a small modification to well-understood surface models. We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements. It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible retro-reflective result.

@@

A microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl,\vechr) \, F(\vecv, \vechr)}{4 \avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of micronormals (NDF), normalized to fulfill $\int_\Omega D(\vech) (\vdot{\vecn}{\vech}) \diff\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$ (where $\etav$, $\etal$ are the IORs of the media $\vecv$, $\vecl$ point to) as selector yields the corresponding BTDF:
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\, \etav^2 \, D(\vecht) \, G_{2}(\vecv,\vecl,\vecht) \, \bigl(1 - F(\vecv, \vecht) \bigr)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retro-reflection.
%Before detailing the impact of such a change, we review some standard properties of microfacet models.

@@

To achieve retro-reflection, in MRM we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
\begin{equation}
  \vecbr(\vecv, \vecl) = \vechr(\vecvv, \vecl) = \frac{\vecvv + \vecl}{\lVert \vecvv + \vecl \rVert}, \text{ with } \vecvv = \text{reflect}(\vecv, \vecn),
\end{equation}
to select the reflecting microfacet. %, we are to re-derive the microfacet model.
For refraction, we can construct a similar refraction back-vector $\vecbt(\vecv, \vecl) = \vecht(\vecvv, \vecl)$.

@@

As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual mirror'' aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$, thus turning glossy forward-reflection into glossy retro-reflection.  \citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface. Though here, there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.

@@

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retro-reflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead. From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}
Thus for both reflection and refraction, the effect on the microsurface BSDF of using the back-vector as selector is to replace $\vecv$ by $\vecvv$ in the Fresnel term and in the denominators.

@@

We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution, and thus replace  $G_{2}(\vecv, \vecl, \vech)$  with $G_{2}(\vecvv, \vecl, \vech)$ in the retro-reflective microfacet model. With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration. For V-cavity masking, taking $\vecvv$ as view direction simply corresponds to selecting $\vecb$ as microfacet normal.
%\todo{\tiny Laurent: I think we could extract the shadowing from Matthias derivation here. The other way to justify the derivation of the full model is to highlight that Smith's shadowing/masking does not depend on the microfacet (appart from the sideness). Hence, whatever the microfacet selector (as long as the sideness matches), the shadowing/masking is the same. Since $\vecv$ and $\vecvv$ share the same slope, we can exchange them in the equation.}

\paragraph*{Final form.}
It follows from the above back-vector modifications that our MRM retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{MRM}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{MRM}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
    f_{t, \text{MRM}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
\begin{figure}[!tb]
  \centering
  \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
  \caption{Comparison of the retro-reflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$. The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$. The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
  \label{fig:brdf_plots}
\end{figure}

To gain some quantitative insight into the behavior of the MRM model, Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue). For both MRM and GGX, the Fresnel factor is set to 1.
\enlargethispage{1\baselineskip}
The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane. In this case. both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retro-reflection direction respectively for GGX and MRM).

@@

We briefly show here that the retro-reflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills reciprocity symmetry and energy conservation requirements.

@@

We first verify that the MRM microfacet model fulfills reciprocity symmetry, i.e. that:
\begin{equation}
  f_{\text{MRM}}(\vecv, \vecl) = f_{\text{MRM}}(\vecl, \vecv) \ , \quad f_{t,\text{MRM}}(\vecv, \vecl) = \frac{\etav^2}{\etal^2}f_{t,\text{MRM}}(\vecl, \vecv) \ .
\end{equation}
For brevity, in analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
We will verify explicitly that the retro-reflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
Note that for this we will need to impose one technical assumption on $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

@@

We now verify that the retro-reflective microfacet model fulfills energy conservation, i.e. that the directional albedos fulfill
\begin{equation}
  \rho_{r, \text{MRM}}(\vecv) + \rho_{t, \text{MRM}}(\vecv) \leq 1.
\end{equation}
For the directional albedos we simply have
\begin{equation}
  \begin{split}
  \rho_{r, \text{MRM}}(\vecv) =
  \int_{\Omega_{\vecv}} f_{r, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_r(\vecvv),\\
  \rho_{t, \text{MRM}}(\vecv) =
  \int_{\Omega_{-\vecv}} f_{t, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_t(\vecvv) \ .
  \end{split}
\end{equation}
That is, the directional albedo of MRM is equivalent to the directional albedo of the regular microfacet BSDF, but evaluated at the reflected direction.
As the original microfacet BSDF fulfills energy conversion, the retro-reflective modification does, too.
%% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
%% \begin{equation}
%%   \begin{split}
%%   \rho(\vecvv) &=
%%   \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
%%   &\leq
%%   \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
%%   = 1.
%%   \end{split}
%% \end{equation}
We further note that under our requirement of a reflection-symmetric microfacet normal distribution, the geometric
configurations for $\vecv$ and $\vecvv$ are identical.
Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retro-reflection are the same, i.e. $\rho_{r, \text{MRM}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{MRM}}(\vecv) = \rho_t(\vecv)$.

@@

Given an implementation of a regular microfacet BSDF, extending it to retro-reflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retro-reflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Kulla17} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables. The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.
\item An implementation may optionally wish to provide a blend between regular and retro-reflective microfacet behavior, to model materials with partial retro-reflective properties (for example this is required in OpenPBR \cite{OpenPBR2024}). This can be achieved simply by blending $f_{r, \text{MRM}}$ and $f_r$, or alternatively via a stochastic selection (e.g. one-sample MIS \cite{Veach1998}) between the original $\vecv$ and reflected $\vecvv$ view directions during sampling.
\end{itemize}

%\noindent
\begin{lstlisting}[
    style=snippet,
    language=GLSL,
    caption=Pseudo code to implement the retro-reflective microfacet model using an existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(Vp, L, N, params);
}

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.99\textwidth]{figures/brdf_theta_L_plots.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{Measured BRDF of retro-reflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
  \label{fig:data_fit}
\end{figure}

>>>>

\teaser{
  \vspace*{-0.5cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/nonretro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/nonretro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/nonretro_shoe_cropped_label.png} \\
    \vspace{0.06cm}
    \includegraphics[width=0.3\linewidth]{figures/cone/retro_cone_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/jacket/retro_jacket_cropped_label.png}
    \includegraphics[width=0.3\linewidth]{figures/shoe/retro_shoe_cropped_label.png}
  \caption{Renders of our ``Minimal Retroreflective Microfacet'' (MRM) retroreflector model (bottom row) obtained by an almost trivial modification of the equivalent roughness GGX microfacet model (top row). The lighting in all cases is aligned with the view direction, exhibiting a strong retroreflective peak in the MRM model.}
   \vspace{-0.5cm}
  \label{fig:teaser}
}

@@

Rendering of retroreflective materials is useful in a variety of contexts. For safety applications such as road markings, signs, vehicles and high-visibility clothing items, retroreflective materials are heavily used (see Figure~\ref{fig:teaser} for some example renders). Realistic rendering of such materials is important for design and visualization purposes, and in a predictive rendering context for example for simulation of visibility in training of autonomous vehicles or for safety certification of road markings.

@@

\paragraph{Previous works.}
Modeling of retroreflectivity in the surface scattering function has received relatively little attention so far. For example, \citet{Guo2017Retro} proposed a model to incorporate the retroreflectivity of prismatic sheets. However, their model only provides pure retroreflection which would lack artistic control. They later extended this model to account for imperfect retroreflection with the addition of a microfacet surface (following Beckmann's normal distribution) above the prismatic sheet~\cite{Guo2018Retro}. However, due to the coupling of retroreflection and rough refraction, they were unable to derive a close form solution for it and relied on fitting of the ground truth model. \citet{Hericz2016} proposed a retroreflective model based on ray tracing simulation of glass bead microstructures, however they also did not provide a closed form BSDF model.

@@

For practical purposes in visual effects, we are interested in a model which is a) visually plausible at least and ideally follows measured observations reasonably well, b) provably energy-conserving and reciprocal, c) efficient and easy to implement, ideally a small modification to well-understood surface models. We present here a model based on the previously published \emph{back-vector} formulation~\cite{BelcourBackvector} which meets all these requirements. It is so simple a modification to a regular microfacet model, that we term it the ``Minimal Retroreflective Microfacet'' (MRM) model which achieves a plausible ive result.

@@

A microfacet BRDF \cite{WalterMicrofacet, HeitzMicrofacet} takes the form (\citet{WalterMicrofacet} Equation (20))
\begin{equation}
  f_r(\vecv, \vecl) = \frac{D(\vechr) \, G_{2}(\vecv,\vecl,\vechr) \, F(\vecv, \vechr)}{4 \avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn}}
  \label{eq:microfacet_brdf}
\end{equation}
where $D$ is the distribution of micronormals (NDF), normalized to fulfill $\int_\Omega D(\vech) (\vdot{\vecn}{\vech}) \diff\vech = 1$, $G_{2}$ is the shadowing-masking function, and $F$ is the Fresnel term. The half-vector $\vechr = \vechr(\vecv, \vecl) = \frac{\vecv + \vecl}{\lVert \vecv + \vecl \rVert}$ is used as the \textit{selector} for microfacets reflecting light from incident direction $\vecl$ to outgoing view direction $\vecv$ (\textit{i.e.} only microfacets with normals aligned with the half-vector will reflect light).
Similarly, using the half-vector of refraction $\vecht = \vecht(\vecv, \vecl) = -\frac{\left(\etav\vecv + \etal\vecl\right)}{\lVert\etav\vecv + \etal\vecl\rVert}$ (where $\etav$, $\etal$ are the IORs of the media $\vecv$, $\vecl$ point to) as selector yields the corresponding BTDF:
\begin{equation}
  f_t(\vecv, \vecl) =
  \frac{\avdot{\vecv}{\vecht}\avdot{\vecl}{\vecht}\, \etav^2 \, D(\vecht) \, G_{2}(\vecv,\vecl,\vecht) \, \bigl(1 - F(\vecv, \vecht) \bigr)}
       {\avdot{\vecv}{\vecn} \avdot{\vecl}{\vecn} \left(\etav \vdot{\vecv}{\vecht} + \etal \vdot{\vecl}{\vecht} \right)^2}
  \label{eq:microfacet_btdf}.
\end{equation}
In Section~4.1 and Section~4.2 of their work, Walter et al.~\cite{WalterMicrofacet} detail the effect of using the half vector as the microfacet selector and obtain Equations~\eqref{eq:microfacet_brdf} and \eqref{eq:microfacet_btdf}.
In this work, we show that using the \emph{back vector}~\cite{BelcourBackvector} instead as a microfacet selector produces a highly plausible model of retroreflection.
%Before detailing the impact of such a change, we review some standard properties of microfacet models.

@@

To achieve retroreflection, in MRM we modify the microfacet model by changing the microfacet normal selector from the half-vector to the \emph{back-vector}~\cite{BelcourBackvector}, defined as
\begin{equation}
  \vecbr(\vecv, \vecl) = \vechr(\vecvv, \vecl) = \frac{\vecvv + \vecl}{\lVert \vecvv + \vecl \rVert}, \text{ with } \vecvv = \text{reflect}(\vecv, \vecn),
\end{equation}
to select the reflecting microfacet. %, we are to re-derive the microfacet model.
For refraction, we can construct a similar refraction back-vector $\vecbt(\vecv, \vecl) = \vecht(\vecvv, \vecl)$.

@@

As illustrated in Figure \ref{fig:backvector}, the reflection $\vecvv$ of the view direction $\vecv$ on the surface can also be interpreted as a reflection of the view direction in a ``virtual mirror'' aligned with $\vecv$ and perpendicular to the surface macronormal $\vecn$, thus turning glossy forward-reflection into glossy retroreflection.  \citet{BelcourBackvector} showed that this reflection is related to the geometry of double scattering on the microsurface. Though here, there is no direct physical interpretation of the back-vector as such a scattering mode, this interpretation can serve as a motivation why the components of the regular microfacet model still work on changing the domain to the back-vector.

@@

\paragraph*{Microfacet reflection term.}
There is experimental evidence~\cite{BelcourBackvector} that using a Fresnel term based on the back-vector can provide a good match to actual retroreflective material measurements for a reflection back-vector based BRDF.
As such, we update the formulation of $\rho(\vecv, \vecm)$~(Equation (11) of \citet{WalterMicrofacet}) to use the reflected view vector $\vecvv$ instead. From the aforementioned equivalence of the Jacobians, it follows that the reflection microsurface BRDF~(Equation (15) of \citet{WalterMicrofacet}) becomes
\begin{equation}
  f_r^m(\vecv, \vecl, \vecm) = F(\vecvv, \vecm) \dfrac{\delta(\vecb_r, \vecm) }{4 \left| \vdot{\vecvv}{\vecb_r} \right|^2}.
\end{equation}
We used the fact that $\left| \vdot{\vecl}{\vecb} \right| = \left| \vdot{\vecvv}{\vecb} \right|$.
Similarly, for the transmissive microsurfacet BTDF (Equation~(18) of \citet{WalterMicrofacet}) we get
\begin{equation}
  f_t^m(\vecv, \vecl, \vecm) = \left(1-F(\vecvv, \vecm)\right) \dfrac{\delta(\vecb_t, \vecm) \eta_o^2 }{\left( \eta_i \vdot{\vecv^\prime}{\vecb_t} + \eta_o \vdot{\vecl}{\vecb_t} \right)^2}.
\end{equation}
Thus for both reflection and refraction, the effect on the microsurface BSDF of using the back-vector as selector is to replace $\vecv$ by $\vecvv$ in the Fresnel term and in the denominators.

@@

We assume Smith or V-cavity masking is applied to the mirrored microfacet distribution, and thus replace  $G_{2}(\vecv, \vecl, \vech)$  with $G_{2}(\vecvv, \vecl, \vech)$ in the retroreflective microfacet model. With Smith masking, this is actually identical to a regular microfacet BSDF, as $\vecvv$ and $\vecv$ share the same geometric configuration. For V-cavity masking, taking $\vecvv$ as view direction simply corresponds to selecting $\vecb$ as microfacet normal.
%\todo{\tiny Laurent: I think we could extract the shadowing from Matthias derivation here. The other way to justify the derivation of the full model is to highlight that Smith's shadowing/masking does not depend on the microfacet (appart from the sideness). Hence, whatever the microfacet selector (as long as the sideness matches), the shadowing/masking is the same. Since $\vecv$ and $\vecvv$ share the same slope, we can exchange them in the equation.}

\paragraph*{Final form.}
It follows from the above back-vector modifications that our MRM retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{MRM}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{MRM}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
    f_{t, \text{MRM}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
\begin{figure}[!tb]
  \centering
  \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (green) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in {45^\circ, 80^\circ}$. The dashed lines are the roughness $0.1$ case, and the dashed lines roughness $0.666$. The grey horizontal line at $1/\pi$ is the Lambert BRDF.}
  \label{fig:brdf_plots}
\end{figure}

To gain some quantitative insight into the behavior of the MRM model, Figure~\ref{fig:brdf_plots} plots the BRDF values of the MRM model (green) compared to the rough diffuse (albedo 1) EON model \cite{Portsmouth2025EON} (orange) and GGX (blue). For both MRM and GGX, the Fresnel factor is set to 1.
\enlargethispage{1\baselineskip}
The $0.1$ roughness case is a very strong, tight peak (note the logarithmic scale on the y-axis), while the $0.666$ roughness case is a much broader peak.
It is evident that MRM BRDF is the same as GGX,with the peak flipped to be back-scattering (at exactly $\theta_o =-\theta_i$, $\phi_o = \phi_i$).
The $\phi_o = 45^\circ$ case demonstrates the situation when the view direction is not aligned in the same azimuthal plane. In this case. both GGX and MRM are dimmer than Lambert or EON, which compensates for the extra energy of the peaks of MRM and GGX (in which almost all of the energy is concentrated, in the reflection or retroreflection direction respectively for GGX and MRM).

@@

We briefly show here that the retroreflective microfacet model of Equation~\eqref{eq:retro_microfacet_model} is plausible, i.e. that it fulfills reciprocity symmetry and energy conservation requirements.

@@

We first verify that the MRM microfacet model fulfills reciprocity symmetry, i.e. that:
\begin{equation}
  f_{\text{MRM}}(\vecv, \vecl) = f_{\text{MRM}}(\vecl, \vecv) \ , \quad f_{t,\text{MRM}}(\vecv, \vecl) = \frac{\etav^2}{\etal^2}f_{t,\text{MRM}}(\vecl, \vecv) \ .
\end{equation}
For brevity, in analogy to the notation for $\vecvv$ we generally let $\vecxx := \text{reflect}(\vecx, \vecn)$.
We will verify explicitly that the retroreflective model is symmetric under the interchange operation $(\vecvv, \vecl) \leftrightarrow (\vecll, \vecv)$, which is the equivalent of the usual reciprocity symmetry $(\vecv, \vecl) \leftrightarrow (\vecl, \vecv)$ for regular microfacet models.
Note that for this we will need to impose one technical assumption on $D$: we require that the microfacet normal distribution is reflection-symmetric around the normal (which is the case for all isotropic distributions and for anisotropic Phong, Beckmann, and GGX).

@@

We now verify that the retroreflective microfacet model fulfills energy conservation, i.e. that the directional albedos fulfill
\begin{equation}
  \rho_{r, \text{MRM}}(\vecv) + \rho_{t, \text{MRM}}(\vecv) \leq 1.
\end{equation}
For the directional albedos we simply have
\begin{equation}
  \begin{split}
  \rho_{r, \text{MRM}}(\vecv) =
  \int_{\Omega_{\vecv}} f_{r, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
  \int_{\Omega_{\vecvv}} f_r(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_r(\vecvv),\\
  \rho_{t, \text{MRM}}(\vecv) =
  \int_{\Omega_{-\vecv}} f_{t, \text{MRM}}(\vecv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl =
  \int_{\Omega_{-\vecvv}} f_t(\vecvv, \vecl) \avdot{\vecl}{\vecn} \diff\vecl = \rho_t(\vecvv) \ .
  \end{split}
\end{equation}
That is, the directional albedo of MRM is equivalent to the directional albedo of the regular microfacet BSDF, but evaluated at the reflected direction.
As the original microfacet BSDF fulfills energy conversion, the retroreflective modification does, too.
%% Which, using equation \eqref{eq:visible} and $\frac{d\vech(\vecvv, \vecl)}{d\vecl} = \frac{1}{4 \vdot{\vecl}{\vech(\vecvv, \vecl)}}= \frac{1}{4 \vdot{\vecvv}{\vech(\vecvv, \vecl)}}$ \cite{WalterMicrofacet}, can be shown to fulfill energy conservation
%% \begin{equation}
%%   \begin{split}
%%   \rho(\vecvv) &=
%%   \int_\Omega \frac{D(\vech(\vecvv, \vecl)) G_{2}(\vecvv,\vecl)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}\vdot{\vecl}{\vecn} d\vecl\\
%%   &\leq
%%   \int_{\lbrace\vech \in \Omega: \vdot{\vecvv}{\vech} \geq 0\rbrace} \frac{D(\vech) G_1(\vecvv,\text{reflect}(\vecvv, \vech))}{\vdot{\vecvv}{\vecn}}\vdot{\vecvv}{\vech} d\vech
%%   = 1.
%%   \end{split}
%% \end{equation}
We further note that under our requirement of a reflection-symmetric microfacet normal distribution, the geometric
configurations for $\vecv$ and $\vecvv$ are identical.
Hence we can conclude $\rho_r(\vecv) = \rho_r(\vecvv)$ and $\rho_t(\vecv) = \rho_t(\vecvv)$, which in turn implies that the albedos of forward and retroreflection are the same, i.e. $\rho_{r, \text{MRM}}(\vecv) = \rho_r(\vecv)$ and $\rho_{t, \text{MRM}}(\vecv) = \rho_t(\vecv)$.

@@

Given an implementation of a regular microfacet BSDF, extending it to retroreflection is extremely straightforward (see the pseudo-code provided in Listing~\ref{lst:implementation}):

\begin{itemize}
\item
  Evaluation merely needs to replace $\vecv$ with $\vecvv$ upfront.
\item
  Similarly, importance sampling of $\vecl$ given $\vecv$ can be realized by replacing $\vecv$ with $\vecvv$ upfront and then importance sampling the regular microfacet BSDF.
  This may include low variance sampling using the domain of visible microfacets \cite{HeitzIS}.
\item
  As the albedos of standard BSDF and retroreflective BSDF are identical, compensating for the energy loss of microfacet single-scattering via a diffuse term \cite{KelemenBRDF, Kulla17} or normalization \cite{TurquinMultipleScattering} can thus be realized using the same data tables. The exact energy compensation of \citet{Heitz2016} can also be applied unchanged.
\item An implementation may optionally wish to provide a blend between regular and retroreflective microfacet behavior, to model materials with partial retroreflective properties (for example this is required in OpenPBR \cite{OpenPBR2024}). This can be achieved simply by blending $f_{r, \text{MRM}}$ and $f_r$, or alternatively via a stochastic selection (e.g. one-sample MIS \cite{Veach1998}) between the original $\vecv$ and reflected $\vecvv$ view directions during sampling.
\end{itemize}

%\noindent
\begin{lstlisting}[
    style=snippet,
    language=GLSL,
    caption=Pseudo code to implement the retroreflective microfacet model using an existing regular microfacet BSDF implementation,
    label=lst:implementation,
    nolol=true,
    frame=trBL
    ]
Sample_result sample_bsdf_retro(vec3 V, vec3 L, vec3 N, Bsdf_params params)
{
    vec3 Vp = -V + 2.0 * dot(V, N) * N;
    return sample_bsdf(Vp, L, N, params);
}

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.99\textwidth]{figures/brdf_theta_L_plots.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{Measured BRDF of retroreflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
  \label{fig:data_fit}
\end{figure}
